---
title: "Exploratory data analysis"
format: html
date: "`r format(Sys.time(), '%d %B, %Y')`"    
toc: true
toc-expand: 3
editor: source
execute: 
  cache: false
editor_options: 
  chunk_output_type: console
knitr:
  opts_chunk: 
    echo: true
    message: false
    warning: false
    code-fold: true
    include: true
    collapse: true
---

```{r}
#| label: setup-env

# Set-up R environment
rm(list=ls())

library(knitr)
library(tidyverse)
library(readxl)
library(DT)
library(arsenal)
library(gtsummary)
library(RColorBrewer)
library(edgeR)
library(plotly)
library(kableExtra)
library(ggpmisc)
library(ggpubr)


# load functions
source("functions.R")

```

```{r}
# load data
file <- "../data/processed/data.RData"
load(file)
```

```{r}
# set graphical theme
mycols <- brewer.pal(8, "Set1")
mycols2 <- brewer.pal(8, "Dark2")

text_size <- 8
mytheme <- theme(legend.title = element_blank(), 
                 legend.text = element_text(size = text_size),
                 panel.grid.major = element_blank(), 
                 panel.grid.minor = element_blank(),
                 axis.text = element_text(size = text_size),
                 axis.title = element_text(size = text_size), 
                 strip.background =element_rect(fill=c("white"), colour = "white"), 
                 strip.text = element_text(colour = 'black', size = text_size + 1))
```


# Introduction

- V2: luteal phase
- V3: follicular phase

# Clinical data

```{r}
#| label: tbl-summary-stats
#| tbl-cap: "Summary statistics of BV, Nugent Score and Age for V2 and V3 overlapping samples."
#| results: asis

# clinical data
df_clin <- df_meta_subset %>%
  left_join(df_meta) 

# extend with delta BV and Nugent scores (V3: folicular phase - V2: luteal phase)
df_clin_ext <- df_clin %>%
  mutate(delta_Nugent = Nugent_v2 - Nugent_v3) %>%
  mutate(delta_BV = as.numeric(BV_v2) - as.numeric(BV_v3)) %>%
  mutate(delta_BV = as.factor(delta_BV))

# clinical data in long format
df_v2 <- df_clin %>%
  dplyr::select(PatID, BV_v2, Nugent_v2, age) %>%
  mutate(ID = paste(PatID, "_V2", sep="")) %>%
  mutate(V = "V2") %>%
  rename(BV = BV_v2, 
         Nugent = Nugent_v2) %>%
  relocate(V, .after = PatID) %>%
  relocate(ID)

df_v3 <- df_clin %>%
  dplyr::select(PatID, BV_v3, Nugent_v3, age) %>%
  mutate(ID = paste(PatID, "_V3", sep="")) %>%
  mutate(V = "V3") %>%
  rename(BV = BV_v3, 
         Nugent = Nugent_v3) %>%
  relocate(V, .after = PatID) %>%
  relocate(ID)

df_clin_long <- rbind(df_v2, df_v3) %>%
  mutate(V = factor(V, levels = c("V3", "V2")))

df_clin_long %>%
  dplyr::select(!c(PatID, ID)) %>%
  tbl_summary(by = V,
              missing_text = "(Missing)") %>%
  as_kable_extra(booktabs = TRUE) %>%
  kable_styling()

```

## Nugent Score

::: {.panel-tabset}

### Histogram
```{r}
#| label: fig-NS-histogram
#| fig-cap: "Histogram of Nugent scores for V2 and V3 samples."

df_clin_long %>%
  ggplot(aes(x=Nugent, fill = V)) + 
  geom_histogram(color = "black", alpha = 0.7) +
  facet_wrap(~V, ncol = 1) + 
  xlab("Nugent score") +
  scale_fill_manual(values = mycols2) + 
  theme_bw() + 
  mytheme

```

### Density
```{r}
#| label: fig-NS-density
#| fig-cap: "Density plot of Nugent scores for V2 and V3 samples."

df_clin_long %>%
  ggplot(aes(x=Nugent, group = V, color = V)) + 
  geom_density(size = 1.2) + 
  scale_color_manual(values = mycols2) + 
  xlab("Nugent score") + 
  theme_bw() + 
  mytheme
```

### Density (change V2 - V3)
```{r}
#| label: fig-NS-density-delta
#| fig-cap: "Distribution of a change of Nugent scores between V3 (follicular phase) and V2 (luteal phase)."

df_clin_ext %>%
  ggplot(aes(x=delta_Nugent)) + 
  geom_histogram(color = "black", fill = mycols[2], alpha = 0.7) +
  scale_color_manual(values = mycols2) + 
  xlab("Nugent score") + 
  theme_bw() + 
  mytheme

```

:::

## BV
```{r}
#| label: tbl-BV-change
#| tbl-cap: "Change of BV category between V3 and V2 phases."
#| results: asis

df_clin_ext %>%
  dplyr::select(delta_BV) %>%
  tbl_summary(missing_text = "(Missing)") %>%
  as_kable_extra(booktabs = TRUE) %>%
  kable_styling()
```

# Normalized count data
```{r}
# calculate PCA on normalized counts
x <- counts_norm_v2 %>%
  column_to_rownames("gene")
pca_v2 <- func_pca(x, df_clin)
head(pca_v2$scores)

x <- counts_norm_v3 %>%
  column_to_rownames("gene")
pca_v3 <- func_pca(x, df_clin)
head(pca_v3$scores)

x <- counts_norm_v2v3 %>%
  column_to_rownames("gene")
pca_v2v3 <- func_pca_v2v3(x, df_clin_long)
head(pca_v2v3$scores)

```

## PCA V2 (luteal phase)

::: {.panel-tabset}

### PC1 vs PC2
```{r}
#| fig-width: 3
#| fig-height: 3
pca_scores <- pca_v2$scores
plot <- func_plot_pca(1, 2, df = pca_scores)
ggplotly(plot)
```

### PC1 vs PC3
```{r}
#| fig-width: 3
#| fig-height: 3
plot <- func_plot_pca(1, 3, df = pca_scores)
ggplotly(plot)
```

### PC2 vs PC3
```{r}
#| fig-width: 3
#| fig-height: 3
plot <- func_plot_pca(2, 3, df = pca_scores)
ggplotly(plot)
```
:::

## Loadings

::: {.panel-tabset}
### PC1
```{r}
ld <- pca_v2$loadings
func_plot_loadings(pc = 1, df_loadings = ld, ntop = 10)
```

### PC2
```{r}
func_plot_loadings(pc = 2, df_loadings = ld, ntop = 10)
```

### PC3
```{r}
func_plot_loadings(pc = 3, df_loadings = ld, ntop = 10)
```
:::

## BV, Nugent score, age
```{r}
# covariates association with PCs
df <- pca_scores
covs <- c("BV_v2", "Nugent_v2", "age")

# initiate empty matrix to store p-values from statistical tests
no_pca <- 5
m_pvalues <- m_stats <-  m_df <-  matrix(data = NA, nrow = length(covs), ncol = no_pca, 
                    dimnames = list(covs, paste("PC", 1:no_pca, sep="")))
colnames(m_pvalues) <- paste("PC", 1:ncol(m_pvalues), sep="")
rownames(m_pvalues) <- covs

# for every covariate and up to selected number of PCA compontents, check for asssociation
for (ivar in 1:length(covs)){
  
  cov <- covs[ivar]
  x <- df %>% pull(cov)
  
  cov_type <- class(x)
  
  if (cov_type == "integer" | cov_type == "numeric"){ # linear regression for numerical variables and PC components
    
    # tmp covariate such as age
    cov <- covs[ivar]
  
    for (j in 1:no_pca){
      
      # regression
      f <- paste(cov, " ~ ", "PC", j, sep="")
      f <- as.formula(f)
      m <- lm(f, data = df)
      m_pvalues[ivar, j] <- summary(m)$coefficient[2, 4]
      
      pc_name <- paste("PC", j, "", sep="")
      pc_value <- df %>% pull(pc_name)
      cortest <- cor.test(pc_value, x)
      m_stats[ivar, j] <- cortest$statistic
      m_df[ivar, j] <- cortest$parameter
     
    }
  }

    if (cov_type == "factor"){ #ANOVA
    
    # tmp covariate such as age
    cov <- covs[ivar]
  
    for (j in 1:no_pca){
      
      # Kruskal-Wallis rank sum test
      pc_name <- paste("PC", j, "", sep="")
      pc_value <- df %>% pull(pc_name)
      
      test <- kruskal.test(pc_value ~ x)
      m_pvalues[ivar, j] <- test$p.value
      
      m_stats[ivar, j] <- test$statistic
      m_df[ivar, j] <- test$parameter
     
    }
  }
}

cols_pink_green <- brewer.pal(6, "PiYG")
cols_spectral <- brewer.pal(6, "Spectral")
col_highlight <- cols_pink_green[2]
df_pvalues <- data.frame(round(m_pvalues, 3))

df_pvalues %>%
  mutate_all(~cell_spec(.x, color = case_when(
  .x <= 0.01 ~ cols_spectral[1],
  .x <= 0.05 & .x > 0.01 ~ cols_spectral[2],
  .default = "black"))) %>%
  kbl(booktabs = T, linesep = "", escape = F) %>%
  kable_paper(full_width = T) %>%
  footnote(general = c("Highlighted in red p-values <= 0.01", "Highlighted in orange p-values <= 0.05"))
```

::: {.panel-tabset}
### PC1 vs PC2 (BV)
```{r}
#| fig-width: 7
#| fig-height: 6
plot <- func_plot_pca_by_cat(1, 2, cov = "BV_v2", pca_scores)
ggplotly(plot)
```
### PC1 vs PC3 (BV)
```{r}
#| fig-width: 7
#| fig-height: 6
plot <- func_plot_pca_by_cat(1, 3, cov = "BV_v2", pca_scores)
ggplotly(plot)
```
### PC2 vs PC3 (BV)
```{r}
#| fig-width: 7
#| fig-height: 6
plot <- func_plot_pca_by_cat(2, 3, cov = "BV_v2", pca_scores)
ggplotly(plot)
```
### PC1 vs PC5 (BV)
```{r}
#| fig-width: 7
#| fig-height: 6
plot <- func_plot_pca_by_cat(1, 5, cov = "BV_v2", pca_scores)
ggplotly(plot)
```
### PC1 vs PC2 (Nugent)
```{r}
#| fig-width: 7
#| fig-height: 6
plot <- func_plot_pca_by_num(1, 2, cov = "Nugent_v2", pca_scores)
ggplotly(plot)
```
### PC1 vs. PC2 (age)
```{r}
#| fig-width: 7
#| fig-height: 6
plot <- func_plot_pca_by_num(1, 2, cov = "age", pca_scores)
ggplotly(plot)
```
### PC1 vs PC5 (age)
```{r}
#| fig-width: 7
#| fig-height: 6
plot <- func_plot_pca_by_num(1, 5, cov = "age", pca_scores)
ggplotly(plot)
```
:::


## PCA V3 (follicular phase)
::: {.panel-tabset}
### PC1 vs PC2
```{r}
#| fig-width: 3
#| fig-height: 3
pca_scores <- pca_v3$scores
plot <- func_plot_pca(1, 2, df = pca_scores)
ggplotly(plot)
```

### PC1 vs PC3
```{r}
#| fig-width: 3
#| fig-height: 3
plot <- func_plot_pca(1, 3, df = pca_scores)
ggplotly(plot)
```

### PC2 vs PC3
```{r}
#| fig-width: 3
#| fig-height: 3
plot <- func_plot_pca(2, 3, df = pca_scores)
ggplotly(plot)
```
:::

## Loadings

::: {.panel-tabset}
### PC1
```{r}
ld <- pca_v3$loadings
func_plot_loadings(pc = 1, df_loadings = ld, ntop = 10)
```

### PC2
```{r}
func_plot_loadings(pc = 2, df_loadings = ld, ntop = 10)
```

### PC3
```{r}
func_plot_loadings(pc = 3, df_loadings = ld, ntop = 10)
```
:::

## BV, Nugent score, age
```{r}
# covariates association with PCs
df <- pca_scores
covs <- c("BV_v3", "Nugent_v3", "age")

# initiate empty matrix to store p-values from statistical tests
no_pca <- 5
m_pvalues <- m_stats <-  m_df <-  matrix(data = NA, nrow = length(covs), ncol = no_pca, 
                    dimnames = list(covs, paste("PC", 1:no_pca, sep="")))
colnames(m_pvalues) <- paste("PC", 1:ncol(m_pvalues), sep="")
rownames(m_pvalues) <- covs

# for every covariate and up to selected number of PCA compontents, check for asssociation
for (ivar in 1:length(covs)){
  
  cov <- covs[ivar]
  x <- df %>% pull(cov)
  
  cov_type <- class(x)
  
  if (cov_type == "integer" | cov_type == "numeric"){ # linear regression for numerical variables and PC components
    
    # tmp covariate such as age
    cov <- covs[ivar]
  
    for (j in 1:no_pca){
      
      # regression
      f <- paste(cov, " ~ ", "PC", j, sep="")
      f <- as.formula(f)
      m <- lm(f, data = df)
      m_pvalues[ivar, j] <- summary(m)$coefficient[2, 4]
      
      pc_name <- paste("PC", j, "", sep="")
      pc_value <- df %>% pull(pc_name)
      cortest <- cor.test(pc_value, x)
      m_stats[ivar, j] <- cortest$statistic
      m_df[ivar, j] <- cortest$parameter
     
    }
  }

    if (cov_type == "factor"){ #ANOVA
    
    # tmp covariate such as age
    cov <- covs[ivar]
  
    for (j in 1:no_pca){
      
      # Kruskal-Wallis rank sum test
      pc_name <- paste("PC", j, "", sep="")
      pc_value <- df %>% pull(pc_name)
      
      test <- kruskal.test(pc_value ~ x)
      m_pvalues[ivar, j] <- test$p.value
      
      m_stats[ivar, j] <- test$statistic
      m_df[ivar, j] <- test$parameter
     
    }
  }
}

cols_pink_green <- brewer.pal(6, "PiYG")
cols_spectral <- brewer.pal(6, "Spectral")
col_highlight <- cols_pink_green[2]
df_pvalues <- data.frame(round(m_pvalues, 3))

df_pvalues %>%
  mutate_all(~cell_spec(.x, color = case_when(
  .x <= 0.01 ~ cols_spectral[1],
  .x <= 0.05 & .x > 0.01 ~ cols_spectral[2],
  .default = "black"))) %>%
  kbl(booktabs = T, linesep = "", escape = F) %>%
  kable_paper(full_width = T) %>%
  footnote(general = c("Highlighted in red p-values <= 0.01", "Highlighted in orange p-values <= 0.05"))
```

::: {.panel-tabset}
### PC1 vs PC2 (BV)
```{r}
#| fig-width: 7
#| fig-height: 6
plot <- func_plot_pca_by_cat(1, 2, cov = "BV_v3", pca_scores)
ggplotly(plot)
```
### PC1 vs PC3 (BV)
```{r}
#| fig-width: 7
#| fig-height: 6
plot <- func_plot_pca_by_cat(1, 3, cov = "BV_v3", pca_scores)
ggplotly(plot)
```
### PC2 vs PC3 (BV)
```{r}
#| fig-width: 7
#| fig-height: 6
plot <- func_plot_pca_by_cat(2, 3, cov = "BV_v3", pca_scores)
ggplotly(plot)
```

### PC1 vs PC4 (BV)
```{r}
#| fig-width: 7
#| fig-height: 6
plot <- func_plot_pca_by_cat(1, 4, cov = "BV_v3", pca_scores)
ggplotly(plot)
```

### PC1 vs PC5 (BV)
```{r}
#| fig-width: 7
#| fig-height: 6
plot <- func_plot_pca_by_cat(1, 5, cov = "BV_v3", pca_scores)
ggplotly(plot)
```

### PC4 vs PC5 (BV)
```{r}
#| fig-width: 7
#| fig-height: 6
plot <- func_plot_pca_by_cat(4, 5, cov = "BV_v3", pca_scores)
ggplotly(plot)
```

### PC1 vs PC2 (Nugent)
```{r}
#| fig-width: 7
#| fig-height: 6
plot <- func_plot_pca_by_num(1, 2, cov = "Nugent_v3", pca_scores)
ggplotly(plot)
```

### PC1 vs PC4 (Nugent)
```{r}
#| fig-width: 7
#| fig-height: 6
plot <- func_plot_pca_by_num(1, 4, cov = "Nugent_v3", pca_scores)
ggplotly(plot)
```

### PC1 vs PC5 (Nugent)
```{r}
#| fig-width: 7
#| fig-height: 6
plot <- func_plot_pca_by_num(1, 5, cov = "Nugent_v3", pca_scores)
ggplotly(plot)
```

### PC1 vs. PC2 (age)
```{r}
#| fig-width: 7
#| fig-height: 6
plot <- func_plot_pca_by_num(1, 2, cov = "age", pca_scores)
ggplotly(plot)
```
### PC1 vs PC5 (age)
```{r}
#| fig-width: 7
#| fig-height: 6
plot <- func_plot_pca_by_num(1, 5, cov = "age", pca_scores)
ggplotly(plot)
```
:::

## PCA V2 & V3 (luteal & follicular)
::: {.panel-tabset}
### PC1 vs PC2
```{r}
#| fig-width: 3
#| fig-height: 3
pca_scores <- pca_v2v3$scores
plot <- func_plot_pca_v2v3(1, 2, df = pca_scores)
ggplotly(plot)
```

### PC1 vs PC3
```{r}
#| fig-width: 3
#| fig-height: 3
plot <- func_plot_pca_v2v3(1, 3, df = pca_scores)
ggplotly(plot)
```

### PC2 vs PC3
```{r}
#| fig-width: 3
#| fig-height: 3
plot <- func_plot_pca_v2v3(2, 3, df = pca_scores)
ggplotly(plot)
```
:::

## Loadings

::: {.panel-tabset}
### PC1
```{r}
ld <- pca_v2v3$loadings
func_plot_loadings(pc = 1, df_loadings = ld, ntop = 10)
```

### PC2
```{r}
func_plot_loadings(pc = 2, df_loadings = ld, ntop = 10)
```

### PC3
```{r}
func_plot_loadings(pc = 3, df_loadings = ld, ntop = 10)
```
:::


## RNA-seq run, BV, Nugent score, age
```{r}
# covariates association with PCs
df <- pca_scores
covs <- c("V", "BV", "Nugent", "age")

# initiate empty matrix to store p-values from statistical tests
no_pca <- 5
m_pvalues <- m_stats <-  m_df <-  matrix(data = NA, nrow = length(covs), ncol = no_pca, 
                    dimnames = list(covs, paste("PC", 1:no_pca, sep="")))
colnames(m_pvalues) <- paste("PC", 1:ncol(m_pvalues), sep="")
rownames(m_pvalues) <- covs

# for every covariate and up to selected number of PCA components, check for association
for (ivar in 1:length(covs)){
  
  cov <- covs[ivar]
  x <- df %>% pull(cov)
  
  cov_type <- class(x)
  
  if (cov_type == "integer" | cov_type == "numeric"){ # linear regression for numerical variables and PC components
    
    # tmp covariate such as age
    cov <- covs[ivar]
  
    for (j in 1:no_pca){
      
      # regression
      f <- paste(cov, " ~ ", "PC", j, sep="")
      f <- as.formula(f)
      m <- lm(f, data = df)
      m_pvalues[ivar, j] <- summary(m)$coefficient[2, 4]
      
      pc_name <- paste("PC", j, "", sep="")
      pc_value <- df %>% pull(pc_name)
      cortest <- cor.test(pc_value, x)
      m_stats[ivar, j] <- cortest$statistic
      m_df[ivar, j] <- cortest$parameter
     
    }
  }

    if (cov_type == "factor"){ #ANOVA
    
    # tmp covariate such as age
    cov <- covs[ivar]
  
    for (j in 1:no_pca){
      
      # Kruskal-Wallis rank sum test
      pc_name <- paste("PC", j, "", sep="")
      pc_value <- df %>% pull(pc_name)
      
      test <- kruskal.test(pc_value ~ x)
      m_pvalues[ivar, j] <- test$p.value
      
      m_stats[ivar, j] <- test$statistic
      m_df[ivar, j] <- test$parameter
     
    }
  }
}

cols_pink_green <- brewer.pal(6, "PiYG")
cols_spectral <- brewer.pal(6, "Spectral")
col_highlight <- cols_pink_green[2]
df_pvalues <- data.frame(round(m_pvalues, 3))

df_pvalues %>%
  mutate_all(~cell_spec(.x, color = case_when(
  .x <= 0.01 ~ cols_spectral[1],
  .x <= 0.05 & .x > 0.01 ~ cols_spectral[2],
  .default = "black"))) %>%
  kbl(booktabs = T, linesep = "", escape = F) %>%
  kable_paper(full_width = T) %>%
  footnote(general = c("Highlighted in red p-values <= 0.01", "Highlighted in orange p-values <= 0.05"))
```

::: {.panel-tabset}
### PC1 vs PC2 (RNA-seq run)
```{r}
#| fig-width: 7
#| fig-height: 6

plot <- func_plot_pca_by_cat(1, 2, cov = "V", pca_scores)
ggplotly(plot)
```

### PC1 vs PC3 (RNA-seq run)
```{r}
#| fig-width: 7
#| fig-height: 6

plot <- func_plot_pca_by_cat(1, 3, cov = "V", pca_scores)
ggplotly(plot)
```

### PC1 vs PC4 (RNA-seq run)
```{r}
#| fig-width: 7
#| fig-height: 6

plot <- func_plot_pca_by_cat(1, 4, cov = "V", pca_scores)
ggplotly(plot)
```

### PC1 vs PC5 (RNA-seq run)
```{r}
#| fig-width: 7
#| fig-height: 6

plot <- func_plot_pca_by_cat(1, 5, cov = "V", pca_scores)
ggplotly(plot)
```

### PC2 vs PC3 (RNA-seq run)
```{r}
#| fig-width: 7
#| fig-height: 6

plot <- func_plot_pca_by_cat(2, 3, cov = "V", pca_scores)
ggplotly(plot)
```

### PC1 vs PC2 (BV)
```{r}
#| fig-width: 7
#| fig-height: 6

plot <- func_plot_pca_by_cat(1, 2, cov = "BV", pca_scores)
ggplotly(plot)
```
:::

```{r}

# removing batch effect?

# Combat?
# RUVg?

install.packages("sva")
library(sva)

expr_matrix <- counts_norm_v2v3 %>%
  column_to_rownames("gene")

combat_dat <- ComBat(expr_matrix, batch=df_clin_long$V, par.prior=TRUE, prior.plots=FALSE)

counts_combat <- combat_dat %>%
  as_tibble(rownames = "gene")


x <- counts_combat %>%
  column_to_rownames("gene")

pca_combat <- func_pca_v2v3(x, df_clin_long)
head(pca_combat$scores)

```


### PC1 vs PC2
```{r}
#| fig-width: 3
#| fig-height: 3
pca_scores <- pca_combat$scores
plot <- func_plot_pca_v2v3(1, 2, df = pca_scores)
ggplotly(plot)


plot <- func_plot_pca_by_cat(1, 2, cov = "BV")
ggplotly(plot)
```

```{r}
# covariates association with PCs
df <- pca_scores
covs <- c("V", "BV", "Nugent", "age")

# initiate empty matrix to store p-values from statistical tests
no_pca <- 5
m_pvalues <- m_stats <-  m_df <-  matrix(data = NA, nrow = length(covs), ncol = no_pca, 
                    dimnames = list(covs, paste("PC", 1:no_pca, sep="")))
colnames(m_pvalues) <- paste("PC", 1:ncol(m_pvalues), sep="")
rownames(m_pvalues) <- covs

# for every covariate and up to selected number of PCA components, check for association
for (ivar in 1:length(covs)){
  
  cov <- covs[ivar]
  x <- df %>% pull(cov)
  
  cov_type <- class(x)
  
  if (cov_type == "integer" | cov_type == "numeric"){ # linear regression for numerical variables and PC components
    
    # tmp covariate such as age
    cov <- covs[ivar]
  
    for (j in 1:no_pca){
      
      # regression
      f <- paste(cov, " ~ ", "PC", j, sep="")
      f <- as.formula(f)
      m <- lm(f, data = df)
      m_pvalues[ivar, j] <- summary(m)$coefficient[2, 4]
      
      pc_name <- paste("PC", j, "", sep="")
      pc_value <- df %>% pull(pc_name)
      cortest <- cor.test(pc_value, x)
      m_stats[ivar, j] <- cortest$statistic
      m_df[ivar, j] <- cortest$parameter
     
    }
  }

    if (cov_type == "factor"){ #ANOVA
    
    # tmp covariate such as age
    cov <- covs[ivar]
  
    for (j in 1:no_pca){
      
      # Kruskal-Wallis rank sum test
      pc_name <- paste("PC", j, "", sep="")
      pc_value <- df %>% pull(pc_name)
      
      test <- kruskal.test(pc_value ~ x)
      m_pvalues[ivar, j] <- test$p.value
      
      m_stats[ivar, j] <- test$statistic
      m_df[ivar, j] <- test$parameter
     
    }
  }
}

cols_pink_green <- brewer.pal(6, "PiYG")
cols_spectral <- brewer.pal(6, "Spectral")
col_highlight <- cols_pink_green[2]
df_pvalues <- data.frame(round(m_pvalues, 3))

df_pvalues %>%
  mutate_all(~cell_spec(.x, color = case_when(
  .x <= 0.01 ~ cols_spectral[1],
  .x <= 0.05 & .x > 0.01 ~ cols_spectral[2],
  .default = "black"))) %>%
  kbl(booktabs = T, linesep = "", escape = F) %>%
  kable_paper(full_width = T) %>%
  footnote(general = c("Highlighted in red p-values <= 0.01", "Highlighted in orange p-values <= 0.05"))
```