---
title: "Heatmaps"
format: html
date: "`r format(Sys.time(), '%d %B, %Y')`"    
toc: true
toc-expand: 3
editor: source
execute: 
  cache: false
editor_options: 
  chunk_output_type: console
knitr:
  opts_chunk: 
    echo: true
    message: false
    warning: false
    code-fold: true
    include: true
    collapse: true
---

```{r}
#| label: setup-env

# Set-up R environment
rm(list=ls())

library(knitr)
library(tidyverse)
library(readxl)
library(DT)
library(arsenal)
library(gtsummary)
library(RColorBrewer)
library(edgeR)
library(plotly)
library(kableExtra)
library(ggpmisc)
library(ggpubr)
library(ggvenn)
library(ggVennDiagram)
library(ggthemes)
library(scales)

# load functions
source("functions.R")
```


```{r}

# define script name
script_name <- "model-compare-BV-groups-in-V03-only-and-check-DMPAs"

# prepare directory to export result tables
dir_exports <- paste("../results/", script_name, "/", sep="")
if (!dir.exists(dir_exports)){
  dir.create(dir_exports)
}


```

# Introduction

The aim of this workbook is to compare BV groups, BV=2 (BV) vs. BV=0 (Normal). 

- no accounting for any covariates (baseline, done previously) (model I)
- adjusting for DMPAs (contraceptives) (model II)
- and after excluding samples with DMPAs (model III)

# Set graphical theme
```{r}
# set graphical theme
mycols <- brewer.pal(8, "Set1")
mycols2 <- brewer.pal(8, "Dark2")

text_size <- 12
mytheme <- theme(legend.title = element_blank(), 
                 legend.text = element_text(size = text_size),
                 panel.grid.major = element_blank(), 
                 panel.grid.minor = element_blank(),
                 axis.text = element_text(size = text_size),
                 axis.title = element_text(size = text_size), 
                 strip.background =element_rect(fill=c("white"), colour = "white"), 
                 strip.text = element_text(colour = 'black', size = text_size + 1))
```

# Load data
```{r}
# load data
file <- "../data/processed/data.RData"
load(file)
rm(counts_norm_v2, counts_norm_v3, counts_norm_v2v3)

# convert counts to matrix (V3)
m_counts_v3 <- counts_v3 %>%
  select(!c("Chr", "Start", "End", "Strand", "Length")) %>%
  column_to_rownames("Geneid") %>%
  as.matrix() 

# convert counts to matrix, remove HIVpos (V2) and samples with missing age
samples_flr <- df_meta %>%
  filter(HIVstatus == "pos" | is.na(age) == TRUE) %>%
  pull(PatID)

# create corresponding meta data
df_meta_v3 <- df_meta %>%
  filter(PatID %in% colnames(m_counts_v3))

```

```{r}
#| label: add-DMPAs-data

# load in meta data
data_contraception <- read_excel("../data/box/KEY_V3_BEA19P035_KB_samples-modified-filled-in-missing-PatID3943.xlsx", sheet = "samples") %>%
  dplyr::select(File, Contraception)

# clincial data for v3
df_clin_v3 <- df_meta_v3 %>%
  rename(BV = BV_v3, 
         Nugent = Nugent_v2) %>%
  left_join(data_contraception, by = c("countfile_v3" = "File"))

```

# Summary of clinical data
```{r}
#| label: tbl-summary-v3
#| tbl-cap: "Summary of clinical data for V3, including DMPA"

df_clin_v3 %>%
  dplyr::select(c(BV, Nugent, age, HIVstatus, Contraception)) %>% 
  tbl_summary(
              missing_text = "(Missing)") %>%
  as_kable_extra(booktabs = TRUE) %>%
  kable_styling()

```

```{r}
#| label: tbl-summary-v3-BV-vs-DMPA
#| tbl-cap: "Summary of clinical data for V3, including DMPA, grouped by BV status"

df_clin_v3 %>%
  dplyr::select(c(BV, Contraception)) %>% 
  tbl_summary(by = BV,
              missing_text = "(Missing)") %>%
  as_kable_extra(booktabs = TRUE) %>%
  kable_styling()
```


# Model I

*What are the differences between BV groups (e.g., BV2 vs BV0), on average, after accounting for age?*

```{r}
#| label: tbl-degs-v3
#| tbl-cap: "Number of differentially down- and up-regulated transcripts, adjusted p-value < 0.05, no lfc threshold."

# input data
matrix_counts <- m_counts_v3
data_meta <- df_meta_v3 %>%
  rename(BV = BV_v3) 

# create edgeR object
BV <- data_meta$BV
age <- data_meta$age
data_cds <- DGEList(matrix_counts, group = BV)

# design matrix
design_matrix <- model.matrix(~0 + BV + age)

# filter lowly expressed genes
keep <- filterByExpr(data_cds, design = design_matrix, group = BV)
#print(summary(keep))
data_cds_flr <- data_cds[keep,,keep.lib.sizes=FALSE]

# extract unfiltered and filtered CPM
x_cpm <- cpm(data_cds, log=TRUE, prior.count = 2)
x_cpm_flr <- cpm(data_cds_flr, log = TRUE, prior.count = 2)

# plot densities
nsamples <- ncol(x_cpm)
cols <- brewer.pal(9, "BuGn")

# Voom transform
data_voom <- voom(data_cds_flr$counts, design_matrix, plot = F)
data_voom <- data_voom$E

# define contrast
my_contrasts <- makeContrasts(BV2_vs_BV0 = BV2 - BV0,
                              BV1_vs_BV0 = BV1 - BV0,
                              BV2_vs_BV1 = BV2 - BV1,
                              levels = design_matrix)
# fit model
fit <- lmFit(data_voom, design_matrix)
qlf <- contrasts.fit(fit, contrast=my_contrasts)
qlf <-  eBayes(qlf)

# combine group results
lfc = 0
decideTests <- decideTests(qlf, lfc = lfc)
decideTests_unadj <- decideTests(qlf, lfc = lfc, adjust.method = "none")

# group results
topTab_B2_B0 <- topTable(qlf, coef = 1, n = Inf) %>%
  as_tibble(rownames = "ENSG")

topTab_B1_B0 <- topTable(qlf, coef = 2, n = Inf) %>%
  as_tibble(rownames = "ENSG")

topTab_B2_B1 <- topTable(qlf, coef = 3, n = Inf) %>%
  as_tibble(rownames = "ENSG") %>%
  rename(logFC.B2_vs_B1 = logFC, 
         AveExpr.B2_vs_B1 = AveExpr,
         t.B2_vs_B1 = t, 
         P.Value.B2_vs_B1 = P.Value, 
         adj.P.Val.B2_vs_B1 = adj.P.Val, 
         B.B2_vs_B1 = B)

# put together
topTab <- topTab_B2_B0 %>%
  left_join(topTab_B1_B0, suffix = c(".B2_vs_B0", ".B1_vs_B0"), by = c("ENSG")) %>% 
  left_join(topTab_B2_B1, suffix = c("", ".B2_vs_B1"), by = c("ENSG"))

# data indicating whether transcripts are DEGs
degs_adj <- data.frame(ENSG = rownames(decideTests), DEGs = decideTests)
degs_unadj <- data.frame(ENSG = rownames(decideTests_unadj), DEGs = decideTests_unadj)

degs <- degs_adj %>%
  left_join(degs_unadj, suffix = c("", ".unadj"), by = c("ENSG"))

# combine and add gene annotations
res <- degs %>%
  left_join(topTab, by = c("ENSG")) %>%
  mutate_at(vars(contains("DEGs")), ~as.factor(.)) %>%
  left_join(counts_anno, by = c("ENSG" = "ensembl_gene_id"))

# DEGs summary
summary(decideTests) %>%
  kbl() %>%
  kable_styling()

res_model_1 <- res

```


## Boxplot of top hits

::: {.panel-tabset}

### BV2 vs. BV0
```{r}
#| label: fig-bv2-bv0
#| fig-cap: "Expression of the top transcripts for a given comparsions, based on the smallest p.value."
#| fig-width: 12
#| fig-height: 12

n <- 9

top_hits <- res %>%
  filter(DEGs.BV2_vs_BV0 != 0) %>%
  arrange(adj.P.Val.B2_vs_B0) %>%
  slice(1:n) %>% 
  pull(ENSG) 

df_top_hits <- data.frame(ENSG = top_hits) %>%
  left_join(res[, c("ENSG", "hgnc_symbol", "gene", "gene_biotype")], by = c("ENSG")) %>%
  mutate(gene = fct_inorder(gene))
  
data_plot <- data_voom[df_top_hits$ENSG, ] %>%
  t() %>% 
  as_tibble(rownames = "PatID") %>%
  pivot_longer(!PatID, names_to = "ENSG", values_to = "CPM") %>%
  left_join(data_meta[, c("PatID", "BV")]) %>%
  left_join(df_top_hits, by = c("ENSG")) 

data_plot %>%
  ggplot(aes(x = BV, y = CPM, fill = BV)) + 
  geom_boxplot(alpha = 0.5) + 
  facet_wrap(~gene) + 
  theme_bw() + 
  mytheme + 
  ylab("log CPM") + 
  theme(legend.position = "none")

```

### BV2 vs. BV0 (logFC)
```{r}
#| label: fig-lfc-bv2-bv0
#| fig-cap: "Expression of the top transcripts for a given comparsions, based on the largest absolute logFC."
#| fig-width: 12
#| fig-height: 12

n <- 9

top_hits <- res %>%
  filter(DEGs.BV2_vs_BV0 != 0) %>%
  arrange(desc(abs(logFC.B2_vs_B0))) %>%
  slice(1:n) %>% 
  pull(ENSG) 

df_top_hits <- data.frame(ENSG = top_hits) %>%
  left_join(res[, c("ENSG", "hgnc_symbol", "gene", "gene_biotype")], by = c("ENSG")) %>%
  mutate(gene = fct_inorder(gene))
  
data_plot <- data_voom[df_top_hits$ENSG, ] %>%
  t() %>% 
  as_tibble(rownames = "PatID") %>%
  pivot_longer(!PatID, names_to = "ENSG", values_to = "CPM") %>%
  left_join(data_meta[, c("PatID", "BV")]) %>%
  left_join(df_top_hits, by = c("ENSG")) 

data_plot %>%
  ggplot(aes(x = BV, y = CPM, fill = BV)) + 
  geom_boxplot(alpha = 0.5) + 
  facet_wrap(~gene) + 
  theme_bw() + 
  mytheme + 
  ylab("log CPM") + 
  theme(legend.position = "none")

```

:::

# Model II

*What are the differences between BV groups (e.g., BV2 vs BV0), on average, after accounting for differences in DMPA and age?*

```{r}
#| label: tbl-degs-v3-model2
#| tbl-cap: "Number of differentially down- and up-regulated transcripts, adjusted p-value < 0.05, no lfc threshold."

# input data
matrix_counts <- m_counts_v3

data_meta <- df_meta_v3 %>%
  rename(BV = BV_v3) %>%
  left_join(data_contraception, by = c("countfile_v3" = "File")) %>% 
  mutate(DMPA = as.factor(Contraception)) %>%
  mutate(DMPA = fct_recode(Contraception, 
                           "No" = "no HC",
                           "Yes" = "DMPA")) %>%
  mutate(DMPA = fct_relevel(DMPA, "No")) 

# create edgeR object
BV <- data_meta$BV
age <- data_meta$age
DMPA <- data_meta$DMPA

data_cds <- DGEList(matrix_counts, group = BV)

# design matrix
design_matrix <- model.matrix(~0 + BV + age + DMPA)

# filter lowly expressed genes
keep <- filterByExpr(data_cds, design = design_matrix, group = BV)
#print(summary(keep))
data_cds_flr <- data_cds[keep,,keep.lib.sizes=FALSE]

# extract unfiltered and filtered CPM
x_cpm <- cpm(data_cds, log=TRUE, prior.count = 2)
x_cpm_flr <- cpm(data_cds_flr, log = TRUE, prior.count = 2)

# Voom transform
data_voom <- voom(data_cds_flr$counts, design_matrix, plot = F)
data_voom <- data_voom$E

# define contrast
my_contrasts <- makeContrasts(BV2_vs_BV0 = BV2 - BV0,
                              BV1_vs_BV0 = BV1 - BV0,
                              BV2_vs_BV1 = BV2 - BV1,
                              levels = design_matrix)
# fit model
fit <- lmFit(data_voom, design_matrix)
qlf <- contrasts.fit(fit, contrast=my_contrasts)
qlf <-  eBayes(qlf)

# combine group results
lfc = 0
decideTests <- decideTests(qlf, lfc = lfc)
decideTests_unadj <- decideTests(qlf, lfc = lfc, adjust.method = "none")

# group results
topTab_B2_B0 <- topTable(qlf, coef = 1, n = Inf) %>%
  as_tibble(rownames = "ENSG")

topTab_B1_B0 <- topTable(qlf, coef = 2, n = Inf) %>%
  as_tibble(rownames = "ENSG")

topTab_B2_B1 <- topTable(qlf, coef = 3, n = Inf) %>%
  as_tibble(rownames = "ENSG") %>%
  rename(logFC.B2_vs_B1 = logFC, 
         AveExpr.B2_vs_B1 = AveExpr,
         t.B2_vs_B1 = t, 
         P.Value.B2_vs_B1 = P.Value, 
         adj.P.Val.B2_vs_B1 = adj.P.Val, 
         B.B2_vs_B1 = B)

# put together
topTab <- topTab_B2_B0 %>%
  left_join(topTab_B1_B0, suffix = c(".B2_vs_B0", ".B1_vs_B0"), by = c("ENSG")) %>% 
  left_join(topTab_B2_B1, suffix = c("", ".B2_vs_B1"), by = c("ENSG"))

# data indicating whether transcripts are DEGs
degs_adj <- data.frame(ENSG = rownames(decideTests), DEGs = decideTests)
degs_unadj <- data.frame(ENSG = rownames(decideTests_unadj), DEGs = decideTests_unadj)

degs <- degs_adj %>%
  left_join(degs_unadj, suffix = c("", ".unadj"), by = c("ENSG"))

# combine and add gene annotations
res <- degs %>%
  left_join(topTab, by = c("ENSG")) %>%
  mutate_at(vars(contains("DEGs")), ~as.factor(.)) %>%
  left_join(counts_anno, by = c("ENSG" = "ensembl_gene_id"))

# DEGs summary
summary(decideTests) %>%
  kbl() %>%
  kable_styling()

res_model_2 <- res

```

## Boxplot of top hits

::: {.panel-tabset}

### BV2 vs. BV0
```{r}
#| label: fig-bv2-bv0-DMPA
#| fig-cap: "Expression of the top transcripts for a given comparsions, based on the smallest p.value."
#| fig-width: 12
#| fig-height: 12

n <- 9

top_hits <- res %>%
  filter(DEGs.BV2_vs_BV0 != 0) %>%
  arrange(adj.P.Val.B2_vs_B0) %>%
  slice(1:n) %>% 
  pull(ENSG) 

df_top_hits <- data.frame(ENSG = top_hits) %>%
  left_join(res[, c("ENSG", "hgnc_symbol", "gene", "gene_biotype")], by = c("ENSG")) %>%
  mutate(gene = fct_inorder(gene))
  
data_plot <- data_voom[df_top_hits$ENSG, ] %>%
  t() %>% 
  as_tibble(rownames = "PatID") %>%
  pivot_longer(!PatID, names_to = "ENSG", values_to = "CPM") %>%
  left_join(data_meta[, c("PatID", "BV")]) %>%
  left_join(df_top_hits, by = c("ENSG")) 

data_plot %>%
  ggplot(aes(x = BV, y = CPM, fill = BV)) + 
  geom_boxplot(alpha = 0.5) + 
  facet_wrap(~gene) + 
  theme_bw() + 
  mytheme + 
  ylab("log CPM") + 
  theme(legend.position = "none")

```

### BV2 vs. BV0 (logFC)
```{r}
#| label: fig-lfc-bv2-bv0-DMPA
#| fig-cap: "Expression of the top transcripts for a given comparsions, based on the largest absolute logFC."
#| fig-width: 12
#| fig-height: 12

n <- 9

top_hits <- res %>%
  filter(DEGs.BV2_vs_BV0 != 0) %>%
  arrange(desc(abs(logFC.B2_vs_B0))) %>%
  slice(1:n) %>% 
  pull(ENSG) 

df_top_hits <- data.frame(ENSG = top_hits) %>%
  left_join(res[, c("ENSG", "hgnc_symbol", "gene", "gene_biotype")], by = c("ENSG")) %>%
  mutate(gene = fct_inorder(gene))
  
data_plot <- data_voom[df_top_hits$ENSG, ] %>%
  t() %>% 
  as_tibble(rownames = "PatID") %>%
  pivot_longer(!PatID, names_to = "ENSG", values_to = "CPM") %>%
  left_join(data_meta[, c("PatID", "BV")]) %>%
  left_join(df_top_hits, by = c("ENSG")) 

data_plot %>%
  ggplot(aes(x = BV, y = CPM, fill = BV)) + 
  geom_boxplot(alpha = 0.5) + 
  facet_wrap(~gene) + 
  theme_bw() + 
  mytheme + 
  ylab("log CPM") + 
  theme(legend.position = "none")

```

:::

# Heatmaps

```{r}
library(ComplexHeatmap)
library(circlize)

# Data Voom

# define function to calculate z scores
func_cal_z_score <- function(x){
  (x - mean(x)) / sd(x)
}
```


## V01

Genes based on Model II (BV2 vs. BV0). 

```{r}
#| fig-width: 10
#| fig-height: 10

# select model
res <- res_model_2

# select genes
genes <- res %>%
  filter(DEGs.BV2_vs_BV0 != 0) %>%
  arrange(adj.P.Val.B2_vs_B0) %>%
  pull(ENSG)

#genes <- genes[1:200]

# input data
x <- data_voom[genes, ]
y <- data_meta %>%
  mutate(group = paste("BV", BV, sep = "")) %>%
  mutate(group = as.factor(group))

print(dim(x))

# scale data
data_zscore <- t(apply(x, 1, func_cal_z_score))
data_heatmap <- data_zscore

# set color scale
print(range(data_heatmap))
cols_wong <- c("#0072b2", "#e69f00")
col_fun = colorRamp2(c(-4, 0, 4), c(cols_wong[1], "white", cols_wong[2]))

# annotations: columns
ha_col <- HeatmapAnnotation(DMPA = data_meta$DMPA,
                            col = list(DMPA = c("No" = "skyblue", "Yes" = "pink")))
set.seed(1234)
ht <- Heatmap(data_heatmap,
        col = col_fun,
        name = "h1",
        show_row_names = FALSE,
        show_column_names = FALSE,
        show_row_dend = FALSE,
        show_column_dend = TRUE,
        column_split = y$group,
        row_title_rot = 0,
        top_annotation = ha_col,
        cluster_column_slices = FALSE,
        use_raster = TRUE)

ht <- draw(ht)

```


## V02

Genes based on Model II (BV2 vs. BV0). Top hits.

```{r}
#| fig-width: 10
#| fig-height: 10

# select model
res <- res_model_2

# select genes
lfc <- 1
genes <- res %>%
  filter(DEGs.BV2_vs_BV0 != 0) %>%
  filter(abs(logFC.B2_vs_B0) >= lfc) %>%
  group_by(DEGs.BV2_vs_BV0) %>%
  arrange(abs(logFC.B2_vs_B0)) %>%
  slice(1:20) %>%
  pull(ENSG)

#genes <- genes[1:200]

# input data
x <- data_voom[genes, ]

idx <- match(rownames(x), counts_anno$ensembl_gene_id)
rows <- counts_anno$hgnc_symbol[idx]
data_rows <- data.frame(ENSG = rownames(x), gene = rows) %>%
  mutate(name = ifelse(gene == "", ENSG, gene))
rownames(x) <- data_rows$name
  
y <- data_meta %>%
  mutate(group = paste("BV", BV, sep = "")) %>%
  mutate(group = as.factor(group))

# scale data
data_zscore <- t(apply(x, 1, func_cal_z_score))
data_heatmap <- data_zscore

# set color scale
print(range(data_heatmap))
cols_wong <- c("#0072b2", "#e69f00")
col_fun = colorRamp2(c(-4, 0, 4), c(cols_wong[1], "white", cols_wong[2]))

# annotations: columns
ha_col <- HeatmapAnnotation(DMPA = data_meta$DMPA,
                            col = list(DMPA = c("No" = "skyblue", "Yes" = "pink")))
set.seed(1234)
ht <- Heatmap(data_heatmap,
        col = col_fun,
        name = "h1",
        show_row_names = TRUE,
        row_names_side = "left",
        show_column_names = FALSE,
        show_row_dend = FALSE,
        show_column_dend = TRUE,
        column_split = y$group,
        row_title_rot = 0,
        top_annotation = ha_col,
        cluster_column_slices = FALSE,
        use_raster = TRUE)

ht <- draw(ht)

```

## V03

Genes based on Model II (BV2 vs. BV0), gene expression values adjusted for age and DMPA

```{r}
#| fig-width: 10
#| fig-height: 10

# input data
y <- data_meta %>%
  mutate(group = paste("BV", BV, sep = "")) %>%
  mutate(group = as.factor(group))

x <- data_voom

# adjust data for age and DMPA
head(fit$coefficients)

coeff_age <- fit$coefficients[, "age"]
coeff_DMPAYes <- fit$coefficients[, "DMPAYes"]

x_age <- matrix(y$age, nrow = nrow(x), ncol = ncol(x), byrow = TRUE)
x_DMPAYes <- matrix(data = as.numeric(y$DMPA)-1, nrow = nrow(x), ncol = ncol(x), byrow = TRUE)

x_adj <- data_voom - 
  coeff_age*x_age -
  coeff_DMPAYes*x_DMPAYes
  
# select model
res <- res_model_2

# select genes
genes <- res %>%
  filter(DEGs.BV2_vs_BV0 != 0) %>%
  arrange(adj.P.Val.B2_vs_B0) %>%
  pull(ENSG)

genes <- genes[1:200]

# subset data
x <- x_adj[genes, ]
print(dim(x))

# scale data
data_zscore <- t(apply(x, 1, func_cal_z_score))
data_heatmap <- data_zscore

# set color scale
print(range(data_heatmap))
cols_wong <- c("#0072b2", "#e69f00")
col_fun = colorRamp2(c(-5, 0, 4), c(cols_wong[1], "white", cols_wong[2]))

# annotations: columns
ha_col <- HeatmapAnnotation(DMPA = data_meta$DMPA,
                            col = list(DMPA = c("No" = "skyblue", "Yes" = "pink")))
set.seed(1234)
ht <- Heatmap(data_heatmap,
        col = col_fun,
        name = "h1",
        show_row_names = FALSE,
        show_column_names = FALSE,
        show_row_dend = FALSE,
        show_column_dend = TRUE,
        column_split = y$group,
        row_title_rot = 0,
        top_annotation = ha_col,
        cluster_column_slices = FALSE,
        use_raster = TRUE)

ht <- draw(ht)

```

## V04
Genes based on Model II (BV2 vs. BV0), gene expression values adjusted for age and DMPA; top hits.

```{r}
#| fig-width: 10
#| fig-height: 10

# input data
y <- data_meta %>%
  mutate(group = paste("BV", BV, sep = "")) %>%
  mutate(group = as.factor(group))

x <- data_voom

# adjust data for age and DMPA
head(fit$coefficients)

coeff_age <- fit$coefficients[, "age"]
coeff_DMPAYes <- fit$coefficients[, "DMPAYes"]

x_age <- matrix(y$age, nrow = nrow(x), ncol = ncol(x), byrow = TRUE)
x_DMPAYes <- matrix(data = as.numeric(y$DMPA)-1, nrow = nrow(x), ncol = ncol(x), byrow = TRUE)

x_adj <- data_voom - 
  coeff_age*x_age -
  coeff_DMPAYes*x_DMPAYes
  
# select model
res <- res_model_2

# select genes
lfc <- 1
genes <- res %>%
  filter(DEGs.BV2_vs_BV0 != 0) %>%
  filter(abs(logFC.B2_vs_B0) >= lfc) %>%
  group_by(DEGs.BV2_vs_BV0) %>%
  arrange(abs(logFC.B2_vs_B0)) %>%
  slice(1:20) %>%
  pull(ENSG)

# subset data
x <- x_adj[genes, ]

idx <- match(rownames(x), counts_anno$ensembl_gene_id)
rows <- counts_anno$hgnc_symbol[idx]
data_rows <- data.frame(ENSG = rownames(x), gene = rows) %>%
  mutate(name = ifelse(gene == "", ENSG, gene))
rownames(x) <- data_rows$name
  
print(dim(x))

# scale data
data_zscore <- t(apply(x, 1, func_cal_z_score))
data_heatmap <- data_zscore

# set color scale
print(range(data_heatmap))
cols_wong <- c("#0072b2", "#e69f00")
col_fun = colorRamp2(c(-3.5, 0, 3.5), c(cols_wong[1], "white", cols_wong[2]))

# annotations: columns
ha_col <- HeatmapAnnotation(DMPA = data_meta$DMPA,
                            col = list(DMPA = c("No" = "skyblue", "Yes" = "pink")))
set.seed(1234)
ht <- Heatmap(data_heatmap,
        col = col_fun,
        name = "h1",
        show_row_names = TRUE,
        row_names_side = "left",
        show_column_names = FALSE,
        show_row_dend = FALSE,
        show_column_dend = TRUE,
        column_split = y$group,
        row_title_rot = 0,
        top_annotation = ha_col,
        cluster_column_slices = FALSE,
        use_raster = TRUE)

ht <- draw(ht)

```
