---
title: "Prepare data sets"
format: html
date: "`r format(Sys.time(), '%d %B, %Y')`"    
toc: true
toc-expand: 3
editor: source
editor_options: 
  chunk_output_type: console
knitr:
  opts_chunk: 
    echo: true
    message: false
    warning: false
    code-fold: true
    include: true
    collapse: true
---

```{r}
#| label: setup-env

# Set-up R environment
rm(list=ls())

library(tidyverse)
library(readxl)
library(DT)
require(biomaRt)

```

# Clinical data

```{r}

# load meta data
file <- "../data/box/metadata_ updated FB 20230504_to OD.xlsx"
df <- read_xlsx(file)

cols_v2_v3 <- c("PatIDPx", "PatID", "age", 
             "BV_Diagnosis_v2...14", "Nugent_Score_v2",
             "BV_Diagnosis_v3...16", "Nugent_Score_v3",
             "Conc_Estradiol_v2 pg/ml", "Below_LLD_E2_V2",
             "Conc_Progesteron_v2 ng/ml",   "Below_LLD_P4_V2", 
             "Plasma_S_Estradiol_pg_mL_v3", "Below_LLD_E2_V3",
             "Plasma_S_Prog_ng_mL_v3",   "Below_LLD_P4_V3"
             )

df_clin <- df %>%
  dplyr::select(any_of(cols_v2_v3)) %>%
  rename(BV_v2 = `BV_Diagnosis_v2...14`, 
         Nugent_v2 = `Nugent_Score_v2`, 
         BV_v3 = `BV_Diagnosis_v3...16`, 
         Nugent_v3 = `Nugent_Score_v3`,
         E2_v2 = `Conc_Estradiol_v2 pg/ml`, 
         P4_v2 = `Conc_Progesteron_v2 ng/ml`, 
         below_LLD_E2_v2 = `Below_LLD_E2_V2`, 
         below_LLD_P4_v2 = Below_LLD_P4_V2, 
         E2_v3= `Plasma_S_Estradiol_pg_mL_v3`, 
         P4_v3 = `Plasma_S_Prog_ng_mL_v3`, 
         below_LLD_E2_v3 = `Below_LLD_E2_V3`, 
         below_LLD_P4_v3 = Below_LLD_P4_V3
         ) %>%
  mutate(PatID = as.character(PatID),
         age = as.numeric(age), 
         BV_v2 = as.numeric(BV_v2), 
         Nugent_v2 = as.numeric(Nugent_v2), 
         BV_v3 = as.numeric(BV_v3), 
         Nugent_v3 = as.numeric(Nugent_v3), 
         E2_v2 = as.numeric(E2_v2), 
         P4_v2 = as.numeric(P4_v2), 
         E2_v3 = as.numeric(E2_v3), 
         P4_v3 = as.numeric(P4_v3), 
         below_LLD_E2_v2 = factor(below_LLD_E2_v2, levels = c("no", "yes", NA)), 
         below_LLD_P4_v2 = factor(below_LLD_P4_v2, levels = c("no", "yes", NA)), 
         below_LLD_E2_v3 = factor(below_LLD_E2_v3, levels = c("no", "yes", NA)), 
         below_LLD_P4_v3 = factor(below_LLD_P4_v3, levels = c("no", "yes", NA)), 
         ) 
  
```


```{r}

# How many samples have matching V2 and V3 entries
# and how many with BV complete measurements

df_clin %>%
  dplyr::select(PatID, age, BV_v2, BV_v3) %>%
  dim()

df_clin %>%
  dplyr::select(PatID, age, BV_v2, BV_v3) %>%
  na.omit() %>%
  dim()

#TODO: update below LLD values for E2 and P4 and double-check numbers
df_clin %>%
  #dplyr::select(PatID, age, BV_v2, BV_v3) %>%
  na.omit() %>%
  dim()

#TODO: check corresponding count data, tabulate set for analyses
```



```{r}
# preview relevant clinical data for V2
df_meta_v2 %>%
  head() %>%
  datatable(caption = paste("Preview of the relevant clinical data for V2 samples: ", nrow(df_meta_v2), " samples and ", ncol(df_meta_v2), " variables.", sep="")) %>%
  formatSignif(columns = c("E2", "P4"), digits = 3)

df_meta_v3 %>%
  head() %>%
  datatable(caption = paste("Preview of the relevant clinical data for V3 samples: ", nrow(df_meta_v3), " samples and ", ncol(df_meta_v3), " variables.", sep="")) %>%
  formatSignif(columns = c("E2", "P4"), digits = 3)
```



```{r}
# udate P4 LLD values to 0.05
# email 2023-05-29
# As a follow up on our discussion last week: Let’s use two decimal points for progesterone (P4), that would mean that LLD is 0.05 for both V2 and V3 (previously V3: 0.0503 and V2: 0.045), then we are consistent with the rest of the manuscript

df_meta_v2 <- df_meta_v2 %>%
  mutate(P4 = ifelse(below_LLD_P4 == "yes", 0.05, P4))

df_meta_v3 <- df_meta_v3 %>%
  mutate(P4 = ifelse(below_LLD_P4 == "yes", 0.05, P4))

```

# Count data

```{r}
# load count data
file_counts_v2 <- "../data/raw/BEA22P129_KB/counts/subreadCounts_hg38ens_minus_frag.txt"
file_counts_v3 <- "../data/raw/BEA22P129_KB/counts_BEA19P035_KB/subreadCounts_hg38ens_minus.txt"

df_counts_v2 <- read_tsv(file_counts_v2, comment = "#")
df_counts_v3 <- read_tsv(file_counts_v3, comment = "#")

# check dimension
# print(dim(df_counts_v2))
# print(dim(df_counts_v3))

# check if order of rows is the same (yes, it is)
# summary(rownames(df_counts_v2) == rownames(df_counts_v3))

# load key data to map BEA samples to PatID
file <- "../data/box/KEY_V2_BEA22P129_KB_samples_.xlsx"
df_keys_v2 <- read_xlsx(file)
df_keys_v2 <- df_keys_v2 %>%
  dplyr::select(File, PatID) %>%
  rename(count_file = File) %>%
  mutate(PatID = as.character(PatID))

file <- "../data/box/KEY_V3_BEA19P035_KB_samples-modified-filled-in-missing-PatID3943.xlsx"
df_keys_v3 <- read_xlsx(file)
df_keys_v3 <- df_keys_v3 %>%
  dplyr::select(File, PatID) %>%
  rename(count_file = File) %>%
  mutate(PatID = as.character(PatID))

# add key data to df_meta
df_meta_keys_v2 <- df_meta_v2 %>%
  left_join(df_keys_v2)

df_meta_keys_v3 <- df_meta_v3 %>%
  left_join(df_keys_v3)

# select matching count data to clinical data
cols_geneinfo <- c("Geneid",  "Chr", "Start", "End", "Strand", "Length")
cols_v2 <- c(cols_geneinfo, df_meta_keys_v2$count_file)
cols_v3 <- c(cols_geneinfo, df_meta_keys_v3$count_file)

df_counts_matched_v2 <- df_counts_v2 %>%
  dplyr::select(all_of(cols_v2))

df_counts_matched_v3 <- df_counts_v3 %>%
  dplyr::select(all_of(cols_v3))

# # check dimensions
# dim(df_counts_matched_v2)
# dim(df_counts_matched_v3)

# # check if order of the columns is matching meta file
# colnames(df_counts_matched_v2) == cols_v2
# colnames(df_counts_matched_v3) == cols_v3

# change count column names to PatID (V2)
df_columns_v2 <- data.frame(count_file = cols_v2) 
df_columns_v2 <- df_columns_v2 %>%
  left_join(df_meta_keys_v2) %>%
  dplyr::select(all_of(c("count_file", "PatIDPx", "PatID"))) %>%
  mutate(columns_new = ifelse(is.na(PatID) == TRUE, count_file, PatID)) 

colnames(df_counts_matched_v2) <- df_columns_v2$columns_new

# change count column names to PatID (V3)
df_columns_v3 <- data.frame(count_file = cols_v3) 
df_columns_v3 <- df_columns_v3 %>%
  left_join(df_meta_keys_v3) %>%
  dplyr::select(all_of(c("count_file", "PatIDPx", "PatID"))) %>%
  mutate(columns_new = ifelse(is.na(PatID) == TRUE, count_file, PatID)) 

colnames(df_counts_matched_v3) <- df_columns_v3$columns_new

# preview count data
df_counts_matched_v2 %>%
  dplyr::select(-c("Chr", "Start", "End", "Strand", "Length")) %>%
  dplyr::select(1:11) %>%
  head() %>%
  datatable(caption = paste("Preview of the count data corresponding to V2 samples: ", nrow(df_counts_v2), " genes", sep=""))

df_counts_matched_v3 %>%
  dplyr::select(-c("Chr", "Start", "End", "Strand", "Length")) %>%
  dplyr::select(1:11) %>%
  head() %>%
  datatable(caption = paste("Preview of the count data corresponding to V3 samples: ", nrow(df_counts_v3), " genes", sep=""))

```

# Annotation data

```{r}

## Use Ensembl 
#ensembl = useMart("ensembl")
#listDatasets(ensembl)

# as the main Ensembl portal is currently unavailable, one needs to specify alternative portal e.g. jul2015.archive.ensembl.org
ensemble_mart=useMart(biomart="ENSEMBL_MART_ENSEMBL", dataset="hsapiens_gene_ensembl")#, host = "jul2015.archive.ensembl.org")

## available attributes available (uncomment below)
# head(listAttributes(ensemble_mart))
# ensembl_attributes <- listAttributes(ensemble.mart)

## save attritbutes to the file
# write.table(listAttributes(ensemble.mart), file='../data/processed/listAttributes.txt', sep='\t', quote=F, row.names=F, col.names=F)


## Retrieving data frame for the selected attributes
my_attributes <- c("ensembl_gene_id", "entrezgene_id", "chromosome_name", "start_position",  "end_position", "strand", "hgnc_symbol", "external_gene_name", "description", "gene_biotype")      
df_ensembl <- getBM(attributes=my_attributes, mart=ensemble_mart)

## Add information to the ensemble genes
## Genes to annotate

df_genes <- tibble(ensembl_gene_id = df_counts_v2$Geneid)

df_ensmbl_annotated <- df_genes %>%
  left_join(df_ensembl) %>% 
  distinct(ensembl_gene_id, .keep_all = TRUE) %>%
  mutate(gene = paste(hgnc_symbol, "@", ensembl_gene_id, sep=""))

```

# Save data

```{r}
#| eval: false
#| code-fold: show

# directory
dir_out <- "../data/processed/"

# rename final data sets
df_meta_v2 <- df_meta_keys_v2
df_meta_v3 <- df_meta_keys_v3

df_counts_v2 <- df_counts_matched_v2
df_counts_v3 <- df_counts_matched_v3

# save final data sets to separate .csv files
file_save <- "df_meta_v2.csv"
write_csv(df_meta_v2, file = paste(dir_out, file_save, sep =""))

file_save <- "df_meta_v3.csv"
write_csv(df_meta_v3, file = paste(dir_out, file_save, sep =""))

file_save <- "df_counts_v2.csv"
write_csv(df_counts_v2, file = paste(dir_out, file_save, sep =""))

file_save <- "df_counts_v3.csv"
write_csv(df_counts_v3, file = paste(dir_out, file_save, sep =""))

file_save <- "df_ensembl_annotated.csv"
write_csv(df_ensmbl_annotated, file = paste(dir_out, file_save, sep=""))

# save final data sets to .RData
file_save <- "data.RData"
save(df_meta_v2, df_meta_v3, df_counts_v2, df_counts_v3, df_ensmbl_annotated, file = paste(dir_out, file_save, sep=""))
```

# Conclusions & comments

-   All samples, V2 (n = 58) and V3 (n = 66), were positively identified in among the files containing clinical information.
-   Corresponding raw count data were extracted successfully from the data delivered by BEA.
-   The relevant clinical and count data were saved to separate .csv files as well as together into .RData file.
